---
title: "behavioural_analysis"
author: "RB"
date: "30/09/2021"
output: html_document
---

NOTES FOR NEXT TIME:
probably better to turn the plotting functions into separate scripts 

load packages
```{r}
library(tidyverse)
library(wesanderson)
library(gridExtra)
library(ggpubr)
library(rstan)
library(plotrix)
library(psycho)
source('scripts/reverseFromZero.R')
source('scripts/CrawfordHowell.R')
```

load online or offline data (online = in scanner)
```{r}
online = 1
```

load in the data 
```{r}
# read in the cleaned data, made in script 1.preprocess_2c.rmd
if (online == 1)
{
  real_data <- read.csv('raw_data/clean/real_data_online.csv')
  sim_data <- read.csv('raw_data/clean/sim_data_online.csv')
  parameters <- read.csv('raw_data/clean/parameters_online.csv')
  p_data <- read.csv('raw_data/clean/p_data_online.csv')
} else {
  real_data <- read.csv('raw_data/clean/real_data_offline.csv')
  sim_data <- read.csv('raw_data/clean/sim_data_offline.csv')
  parameters <- read.csv('raw_data/clean/parameters_offline.csv')
  p_data <- read.csv('raw_data/clean/p_data_offline.csv')
}

# read in cleaned questionnaire data
Q_results <- read.csv('results/questionnaire_results.csv')

# read parameters
fmri_vals <- read.csv('raw_data/clean/fmri_vals.csv')

# set variable to data for analysis
if (online == 1){
  p_data['chosen_dev'] <- 0
}

p_data['p'] <- TRUE
real_data ['p'] <- FALSE
sim_data['p'] <- FALSE

## bind the patient data and healthy data
data <- rbind(real_data,p_data)

```


### RT analysis ### 
do reaction times increase at the start of devaluation?
do rt measures correlate with model params?

```{r}
# first let's get the n not including the patient 
# useful for some analyses 
n = length(parameters$lr)

# create RT dataframe with all the states we're interested in
# log transform all individual RTs
RT <- data %>%
  mutate(RT = endDice-startDice) %>%
  #mutate(RT = log(RT)) %>%
  select(state,stateTime,stateName,subject_nr,chosenDice,correctDice,correct,RT) 

# compare RTs for correct vs incorrect choices during extinction 
extinction_correct <- RT %>%
  filter(stateName == 'extinction') %>%
  filter(correct == 1) %>%
  group_by(subject_nr) %>%
  mutate(RTmean=mean(RT)) %>%
  distinct(RTmean)
extinction_correct = extinction_correct$RTmean

extinction_incorrect <- RT %>%
  filter(stateName == 'extinction') %>%
  filter(correct == 0) %>%
  group_by(subject_nr) %>%
  mutate(RTmean=mean(RT)) %>%
  distinct(RTmean)
extinction_incorrect = extinction_incorrect$RTmean

t.test(extinction_correct,extinction_incorrect)
sd(extinction_correct)
sd(extinction_incorrect)

# compare RTs for extinction vs filling
extinction_RT <- RT %>%
  filter(stateName == 'extinction') %>%
  filter(stateTime < 4) %>%
  group_by(subject_nr) %>%
  mutate(RTmean=mean(RT)) %>%
  distinct(RTmean, stateName)
extinction_RT_mean <- extinction_RT$RTmean  

filling_RT <- RT %>%
  filter(stateName == 'fill') %>%
  group_by(subject_nr) %>%
  mutate(RTmean=mean(RT)) %>%
  distinct(RTmean, stateName)
filling_RT_mean <- filling_RT$RTmean 

t.test(extinction_RT_mean, filling_RT_mean)
sd(extinction_RT_mean)
sd(filling_RT_mean)

RT1_set <- rbind(extinction_RT,filling_RT)

rt1_plot <- ggplot(RT1_set, aes(stateName, RTmean)) +
  geom_violin()

## now calculate devalutation effect on RT and correlate with questionnaire measures
rt_diff = extinction_RT_mean - filling_RT_mean

rt_diff <- as.data.frame(rt_diff) %>%
  rowid_to_column() %>%
  mutate(subject_nr = rowid+1) %>%
  select(subject_nr, rt_diff)

Q_results <- merge(Q_results,rt_diff)

# a few explorative correlations
cor.test(Q_results$AMI,Q_results$rt_diff)
cor.test(Q_results$HADS,Q_results$rt_diff)
cor.test(Q_results$GRIT,Q_results$rt_diff)
cor.test(rt_diff$rt_diff[1:n],parameters$tau)
cor.test(rt_diff$rt_diff[1:n],parameters$lr)
cor.test(parameters$lr,parameters$tau)

plotting_params <- as.data.frame(cbind(parameters$lr,parameters$tau)) %>%
  rename(lr=V1,tau=V2)
plotting_params['subj'] <- rep(1:n)

# plot lr and tau 
lr_tau <- ggplot(plotting_params, aes(x=lr, y=tau)) +
  geom_point() +
  theme_classic() +
  scale_x_continuous(n.breaks=5,limits=c(0,1)) +
  theme(legend.position='none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=17),
        plot.title=element_text(size=17)) +
  labs(title='Winning model parameters',x='Learning rate',y='Inverse temperature') 

```


### CHOICE PROBABILITIES DURING DEVALUATION ANALYSIS ###
analyzing whether participants took the devaluation into account in their choices 

Probability of choosing the correct dice on reinstatement states 
correct = dice associated to non devalued token
```{r}
# create reinstatement states data frame
reinstatement_states <- data %>%
  filter(stateName == 'reinstatement') %>%
  select(state, stateName, chosenDice, correct, subject_nr, trialIndex, stateTime, p) %>%
  group_by(subject_nr, p ) 
```

get the mean and sd of percentage chose correct for reinstatement states for each ppt
```{r}
reinstatement_states <- reinstatement_states %>%
  mutate(pptMean = mean(correct)) %>%
  mutate(pptSD = sd(correct))
```

same again but for extinction states
technically no correct answer as reward probabilities are both set to 0 here
but in theory participants should take into account the devaluation into their choices 
```{r}
# isolate the devalued states and select the needed variables 
extinction_states <- data %>% 
  select(subject_nr, stateName, correct, chosenDice, stateTime, state, trialIndex, p) %>%
  group_by(subject_nr, p) %>%
  filter(stateName == 'extinction')

```

calculate the means and sd's for the probability of choosing the correct dice during extinction 
```{r}
extinction_states <- extinction_states %>%
  mutate(pptMean = mean(correct)) %>%
  mutate(pptSD = sd(correct))
```

plot these probabilities as box plots for both states
```{r}
# box plots for p(chose correct) for each ppt
plotting_1 <- rbind(reinstatement_states,extinction_states) %>%
  select(subject_nr,pptMean,stateName,p) %>%
  rename(patient = p) %>%
  distinct(subject_nr,pptMean,stateName,patient)

box_1 <- ggplot(plotting_1, aes(y=pptMean, x=stateName, group=stateName, fill=stateName)) +
  geom_boxplot(data = plotting_1 %>% filter(!patient==TRUE)) +
  geom_jitter(aes(y=pptMean,x=stateName,group=subject_nr, colour=patient),width=0.2, height = 0) +
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 2)) +
  theme(axis.line = element_line(color='black'),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        plot.title=element_text(size=14),
        legend.text=element_text(size=12),
        legend.title = element_text(size=12)) +
  labs(title = "Choice probability for still valuable option during devaluation states", y="p(chose correct)") +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  theme(text = element_text(size=10)) 


# individual histograms for devaluation effect during extinction
hist_extinction <- plotting_1 %>%
  filter(stateName=='extinction') %>%
  mutate(pptMean=pptMean-0.5) # gives devaluation effect

extinction_deval_hist <- ggplot(hist_extinction,aes(x=pptMean)) +
  geom_histogram(binwidth = 0.1,colour='black',fill='#FF9B5E') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Extinction - Devaluation effect',x='Choice prob - Chance',y='No. Participants') +
  scale_x_continuous(breaks=c(-0.1,0,0.1,0.2,0.3,0.4,0.5))
  
```

significance testing
```{r}
# generate 'devaluation effect' on choice probability
# mean choice probability in state - chance level (50%)

# for reinstatemnet first
reinstatement_devaluation <- reinstatement_states %>%
  select(subject_nr,pptMean,pptSD) %>%
  distinct(subject_nr,pptMean,pptSD) %>%
  mutate(devaluation=pptMean-0.5)

# now extinction
extinction_devaluation <- extinction_states %>%
  select(subject_nr,pptMean,pptSD) %>%
  distinct(subject_nr,pptMean,pptSD) %>%
  mutate(devaluation=pptMean-0.5)

t.test(reinstatement_devaluation$devaluation)
t.test(extinction_devaluation$devaluation)

cor.test(reinstatement_devaluation$devaluation[1:n],parameters$lr)
cor.test(reinstatement_devaluation$devaluation[1:n],parameters$tau)
cor.test(extinction_devaluation$devaluation[1:n],parameters$lr)
cor.test(extinction_devaluation$devaluation[1:n],parameters$tau)
cor.test(reinstatement_devaluation$devaluation[1:n],extinction_devaluation$devaluation[1:n])
cor.test(reinstatement_devaluation$devaluation[1:n],rt_diff$rt_diff[1:n])
cor.test(extinction_devaluation$devaluation[1:n],rt_diff$rt_diff[1:n])

plot(reinstatement_devaluation$devaluation[1:n],parameters$tau)
plot(extinction_devaluation$devaluation[1:n],parameters$tau)
plot(reinstatement_devaluation$devaluation[1:n],extinction_devaluation$devaluation[1:n])

# patient vs control 
CrawfordHowell(extinction_devaluation$devaluation[22],extinction_devaluation$devaluation[1:21])
CrawfordHowell(reinstatement_devaluation$devaluation[22],reinstatement_devaluation$devaluation[1:21])
```

### LEARNING CURVE ANLAYSIS ###

Learning
Here we're going to isolate the probability of choosing dice A for each state
very long code as we're doing all states up to state 21
Could probably use a for loop to do a lot of this 

```{r}
## set a few variable specific to each task version
if(online==0){
  # states that need to have their choices flipped for consistent average choice
  s1=9
  s2=10
  s3=11
  s4=12
  # length of states for the plot
  FL <- 10
  EL <- 8
  RL <- 10
  SL1 <- 10
  SL2 <- 8
  PL <- 4
} else if(online==1){
  # states that need to have their choices flipped for consistent average choice
  s1=8
  s2=9
  s3=10
  s4=11
  # length of states for the plot
  FL <- 10
  EL <- 8
  RL <- 12
  SL1 <- 12
  SL2 <- 8
  PL <- 4
}

# first lets run on the real data then simulated
datasets = c('real','sim')
for (j in 1:2){
  # set data to real then sim
  if (datasets[j] == 'real'){
    data <- rbind(real_data,p_data)
  } else if (datasets[j] == 'sim'){
    data <- sim_data
  }
  
  ## then run analysis
  
  # first we need the adapted trial number for the fill states
  # trial 1 of extinction will be trial 0, so need to reverse fill
  fill <- data %>%
    select(stateName,state,stateTime,subject_nr,chosenDice,p) %>%
    filter(stateName == 'fill') %>%
    rename(trialNr = stateTime) %>%
    rowid_to_column()

  # reverse choices for state 8 so consistent for choice prob analysis
  for (i in fill$rowid){
    if(fill[i,]$state == s1 & fill[i,]$chosenDice == 1){
      fill[i,]$chosenDice = 2
    } else if (fill[i,]$state == s1 & fill[i,]$chosenDice == 2){
      fill[i,]$chosenDice = 1
    }
  }

  adaptedTrialNr = reverseFromZero(fill)
  
  fill['adaptedTrialNr'] = adaptedTrialNr
  
  fill <- fill %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)

  ## now the extinction states
  extinction <- data %>%
    select(stateName,stateTime,state,subject_nr,chosenDice,p) %>%
    filter(stateName=='extinction') %>%
    rename(trialNr = stateTime) %>%
    rowid_to_column()

  # reverse choices for state 9 so consistent for choice prob analysis
  for (i in extinction$rowid){
    if(extinction[i,]$state == s2 & extinction[i,]$chosenDice == 1){
      extinction[i,]$chosenDice = 2
    } else if (extinction[i,]$state == s2 & extinction[i,]$chosenDice == 2){
      extinction[i,]$chosenDice = 1
    }
  }

  extinction['adaptedTrialNr'] <- extinction$trialNr-1 # want first trial of extinction to be trial 0
  
  extinction <- extinction %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)
  
  ex_l=max(extinction$adaptedTrialNr) #extinction length 

  ## now the reinstatement states
  reinstatement <- data %>%
    select(stateName,stateTime,state,subject_nr,chosenDice,p) %>%
    filter(stateName=='reinstatement') %>%
    rename(trialNr = stateTime) %>%
    rowid_to_column()
  
  # reverse choices for state 9 so consistent for choice prob analysis
  for (i in reinstatement$rowid){
    if(reinstatement[i,]$state == s3 & reinstatement[i,]$chosenDice == 1){
      reinstatement[i,]$chosenDice = 2
    } else if (reinstatement[i,]$state == s3 & reinstatement[i,]$chosenDice == 2){
      reinstatement[i,]$chosenDice = 1
    }
  }
  
  reinstatement['adaptedTrialNr'] <- reinstatement$trialNr+ex_l
  
  reinstatement <- reinstatement %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)
  


  # switch
  
  switch <- data %>%
    select(stateName,stateTime,subject_nr,chosenDice,p) %>%
    filter(stateName=='switch') %>%
    rename(trialNr = stateTime) %>%
    rowid_to_column()
  
  switch['adaptedTrialNr'] = switch$trialNr+(EL+RL)
  
  switch_1 <- switch %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)
  
  adaptedTrialNr = reverseFromZero(switch)
  switch['adaptedTrialNr'] = adaptedTrialNr

  switch_2 <- switch %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)

  ## now the post state
  post <- data %>%
    select(stateName,stateTime,state,subject_nr,chosenDice,p) %>%
    filter(stateName=='post') %>%
    rename(trialNr = stateTime) %>%
    rowid_to_column()
  
  
  # reverse choices for state 11 so consistent for choice prob analysis
  for (i in post$rowid){
    if(post[i,]$state == s4 & post[i,]$chosenDice == 1){
      post[i,]$chosenDice = 2
    } else if (post[i,]$state == s4 & post[i,]$chosenDice == 2){
      post[i,]$chosenDice = 1
    }
  }
  
  post['adaptedTrialNr'] <- post$trialNr-1
  
  post <- post %>%
    group_by(adaptedTrialNr,subject_nr,p) %>%
    mutate(pChoice = mean(chosenDice)-1) %>%
    distinct(adaptedTrialNr, pChoice, subject_nr,p)
  
  # now lets rename the pChoice variables to real and sim pChoice
  if (datasets[j] == 'real'){
    fill_total <- fill %>%
      rename(pChoiceReal = pChoice)
    extinction_total <- extinction %>%
      rename(pChoiceReal = pChoice)
    reinstatement_total <- reinstatement %>%
      rename(pChoiceReal = pChoice)
    switch_1_total <- switch_1 %>%
      rename(pChoiceReal = pChoice)
    switch_2_total <- switch_2 %>%
      rename(pChoiceReal = pChoice)
    post_total <- post %>%
      rename(pChoiceReal = pChoice)
  } else if (datasets[j] == 'sim'){
    fill <- fill %>%
      rename(pChoiceSim = pChoice)
    extinction <- extinction %>%
      rename(pChoiceSim = pChoice)
    reinstatement <- reinstatement %>%
      rename(pChoiceSim = pChoice)
    switch_1 <- switch_1 %>%
      rename(pChoiceSim = pChoice)
    switch_2 <- switch_2 %>%
      rename(pChoiceSim = pChoice)
    post <- post %>%
      rename(pChoiceSim = pChoice)
  }
}

n = max(data$subject_nr)

# bind by adapted trial number and subject number 
fill <- merge(fill,fill_total,by=c('subject_nr','adaptedTrialNr','p'))
extinction <- merge(extinction,extinction_total,by=c('subject_nr','adaptedTrialNr','p'))
reinstatement <- merge(reinstatement,reinstatement_total,by=c('subject_nr','adaptedTrialNr','p'))
switch_1 <- merge(switch_1,switch_1_total,by=c('subject_nr','adaptedTrialNr','p'))
switch_2 <- merge(switch_2,switch_2_total,by=c('subject_nr','adaptedTrialNr','p'))
post <- merge(post,post_total,by=c('subject_nr','adaptedTrialNr','p'))

## calculate averages and standard errors for each trial within a state across participants 
# first calculate all the averages 

fill_total <- fill %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice) %>%
  filter(adaptedTrialNr > -(FL+1)) # only use first 10 trials for plot (-1:-10)

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(fill_total$pChoiceReal,fill_total$pChoiceSim)
semChoice <- c(fill_total$semChoice,rep(NA,FL))
dataType <- rep('actual',(FL*2))
dataType[(FL+1):(FL*2)] <- 'model'
adaptedTrialNr <- rep(fill_total$adaptedTrialNr,2)

fill_total <- as.data.frame(cbind(meanChoice,semChoice))
fill_total['dataType'] <- dataType
fill_total['adaptedTrialNr'] <- adaptedTrialNr

# now the same for extinction
extinction_total <- extinction %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice)

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(extinction_total$pChoiceReal,extinction_total$pChoiceSim)
semChoice <- c(extinction_total$semChoice,rep(NA,EL))
dataType <- rep('actual',(EL*2))
dataType[(EL+1):(EL*2)] <- 'model'
adaptedTrialNr <- rep(extinction_total$adaptedTrialNr,2)

extinction_total <- as.data.frame(cbind(meanChoice,semChoice))
extinction_total['dataType'] <- dataType
extinction_total['adaptedTrialNr'] <- adaptedTrialNr

# now for reinstatement
reinstatement_total <- reinstatement %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice) %>%
  filter(adaptedTrialNr < (EL+RL))

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(reinstatement_total$pChoiceReal,reinstatement_total$pChoiceSim)
semChoice <- c(reinstatement_total$semChoice,rep(NA,RL))
dataType <- rep('actual',(RL*2))
dataType[(RL+1):(RL*2)] <- 'model'
adaptedTrialNr <- rep(reinstatement_total$adaptedTrialNr,2)

reinstatement_total <- as.data.frame(cbind(meanChoice,semChoice))
reinstatement_total['dataType'] <- dataType
reinstatement_total['adaptedTrialNr'] <- adaptedTrialNr

# switches
switch_1_total <- switch_1 %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice) %>%
  filter(adaptedTrialNr < (EL+RL+SL1+1))

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(switch_1_total$pChoiceReal,switch_1_total$pChoiceSim)
semChoice <- c(switch_1_total$semChoice,rep(NA,SL1))
dataType <- rep('actual',(SL1*2))
dataType[(SL1+1):(SL1*2)] <- 'model'
adaptedTrialNr <- rep(switch_1_total$adaptedTrialNr,2)

switch_1_total <- as.data.frame(cbind(meanChoice,semChoice))
switch_1_total['dataType'] <- dataType
switch_1_total['adaptedTrialNr'] <- adaptedTrialNr

# reinstatement 2
switch_2_total <- switch_2 %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice) %>%
  filter(adaptedTrialNr > -(SL2+1))

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(switch_2_total$pChoiceReal,switch_2_total$pChoiceSim)
semChoice <- c(switch_2_total$semChoice,rep(NA,SL2))
dataType <- rep('actual',(SL2*2))
dataType[(SL2+1):(SL2*2)] <- 'model'
adaptedTrialNr <- rep(switch_2_total$adaptedTrialNr,2)

switch_2_total <- as.data.frame(cbind(meanChoice,semChoice))
switch_2_total['dataType'] <- dataType
switch_2_total['adaptedTrialNr'] <- adaptedTrialNr

# now for post
post_total <- post %>%
  group_by(adaptedTrialNr) %>%
  mutate(semChoice = sd(pChoiceReal)/sqrt(n)) %>%
  mutate(pChoiceReal = mean(pChoiceReal)) %>%
  mutate(pChoiceSim = mean(pChoiceSim)) %>%
  distinct(adaptedTrialNr,pChoiceReal,pChoiceSim,semChoice)

# then get into correct format to plot predicted and actual choice together
meanChoice <- c(post_total$pChoiceReal,post_total$pChoiceSim)
semChoice <- c(post_total$semChoice,rep(NA,PL))
dataType <- rep('actual',(PL*2))
dataType[(PL+1):(PL*2)] <- 'model'
adaptedTrialNr <- rep(post_total$adaptedTrialNr,2)

post_total <- as.data.frame(cbind(meanChoice,semChoice))
post_total['dataType'] <- dataType
post_total['adaptedTrialNr'] <- adaptedTrialNr


## plotting time
reversals_1 <- rbind(fill_total,extinction_total,reinstatement_total,switch_1_total) 

plot_reversal_1_x <- ggplot(reversals_1, aes(x=adaptedTrialNr, y=meanChoice, group=dataType, color = dataType)) +
  geom_line(size=1) + 
  geom_errorbar(aes(ymin=meanChoice-semChoice,ymax=meanChoice+semChoice, color=dataType)) +
  geom_vline(xintercept = c(0,EL,(EL+RL)), linetype = "dotted") +
  geom_vline(xintercept = c(EL+RL+SL1), linetype = "dashed", size=1) +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(-FL,(EL+RL+SL1)), breaks = seq((-FL+2),(EL+RL+SL1),by=4)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text=element_text(size=12),
        axis.title=element_text(size=14),
        plot.title=element_text(size=16)) +
  labs(title = "Choice Probability - Stochastic Verosion", x = "Trial number", y = "Choice probability, non-devalued option") + 
  annotate(
    "text", label = "F",
    x = -4, y = 1.1, size = 5, colour = "red", fontface=2
  ) +
  annotate(
    "text", label = "EX",
    x = 4, y = 1.1, size = 5, colour = "red", fontface=2
  ) +
  annotate(
    "text", label = "R",
    x = (RL+2), y = 1.1, size = 5, colour = "red", fontface=2
  ) +
  annotate(
    "text", label = "S",
    x = (RL+SL1+2), y = 1.1, size = 5, colour = "red", fontface=2
  )


reversals_1_y <- rbind(switch_2_total,post_total)

plot_reversal_1_y <- ggplot(reversals_1_y, aes(x=adaptedTrialNr, y=meanChoice, group=dataType, colour=dataType)) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=meanChoice-semChoice,ymax=meanChoice+semChoice, color=dataType)) +
  geom_vline(xintercept = c(0), linetype = "dotted") +
  geom_vline(xintercept = c(-8), linetype = "dashed", size=1) +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(-SL2,PL), breaks = seq((-SL2),(PL),by=4)) +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14),
        plot.title=element_text(size=18),
        legend.key.size =unit(2,'cm'),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14)) +
  labs(title = "", x = "Trial number", y = "Choice Probability") + 
  annotate(
    "text", label = "EM",
    x = 2, y = 1.1, size = 5, colour = "red", fontface=2
  ) +
  annotate(
    "text", label = "S",
    x = -4, y = 1.1, size = 5, colour = "red", fontface=2
  )

plot1 <- plot_reversal_1_x
plot2 <- plot_reversal_1_y
grid.arrange(plot1, plot2, ncol=2)

```


## brain behaviour correlations ##
```{r}
if (online == 1){
  
  # devaluation effects on choice prob
  # primarily interest in vmpfc_m and vs_rd
  cor.test(reinstatement_devaluation$devaluation[1:21],fmri_vals$vmpfc_cd[1:21])
  cor.test(reinstatement_devaluation$devaluation[1:21],fmri_vals$vmpfc_rd[1:21])
  cor.test(reinstatement_devaluation$devaluation[1:21],fmri_vals$vs_rd[1:21])
  cor.test(reinstatement_devaluation$devaluation[1:21],fmri_vals$vmpfc_m[1:21])
  cor.test(reinstatement_devaluation$devaluation[1:21],fmri_vals$dmpfc_cd[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$vmpfc_cd[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$vmpfc_rd[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$vs_rd[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$vmpfc_m[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$dmpfc_cd[1:21])
  cor.test(extinction_devaluation$devaluation[1:21],fmri_vals$vmpfc_rd[1:21])

  
  ## plotting the significant ones
  MExtDeval <- as.data.frame(cbind(fmri_vals$vmpfc_m,extinction_devaluation$devaluation))
  
  MExtDeval_plot <- ggplot(MExtDeval, aes(x=V1,y=V2)) +
    geom_point() +
    geom_smooth(method='lm',se=F,colour='red') +
    labs(title='Extinction',y="Probability of choosing \n non-devalued option - chance", x='Motivation signal - mPFC') +
    theme_bw() +
    theme(axis.title = element_text(size=18),
          axis.text = element_text(size=18),
          plot.title = element_text(size=18))
  
  MReinDeval <- as.data.frame(cbind(fmri_vals$vmpfc_m,reinstatement_devaluation$devaluation))
  
  MReinDeval_plot <- ggplot(MReinDeval, aes(x=V1,y=V2)) +
    geom_point() +
    geom_smooth(method='lm',se=F,colour='red') +
    labs(title='Reinstatement',y="Probability of choosing \n non-devalued option - chance", x='Motivation signal - mPFC') +
    theme_bw() +
    theme(axis.title = element_text(size=18),
          axis.text = element_text(size=18),
          plot.title = element_text(size=18))
  
  DrwdExtDeval <- as.data.frame(cbind(fmri_vals$vs_rd,extinction_devaluation$devaluation))
  
  DrwdExtDeval_plot <- ggplot(DrwdExtDeval, aes(x=V1,y=V2)) +
    geom_point() +
    geom_smooth(method='lm',se=F,colour='blue') +
    labs(y="Probability of choosing \n non-devalued option - chance", x='Devaluation of Reward Signal - Ventral Striatum') +
    theme_bw() +
    theme(axis.title = element_text(size=18),
          axis.text = element_text(size=18))
    
                          
  # tau, primarily interested in dmPFC_CD - should correlate as used for exploration
  cor.test(parameters$tau,fmri_vals$dmpfc_cd)
  plot(parameters$tau,fmri_vals$dmpfc_cd)
  
  # correlations between signals
  cor.test(fmri_vals$vmpfc_m,fmri_vals$vs_rd)
  cor.test(fmri_vals$vmpfc_m,fmri_vals$vmpfc_rd)
  cor.test(fmri_vals$vmpfc_m,fmri_vals$vmpfc_cd)
  cor.test(fmri_vals$vmpfc_cd,fmri_vals$vs_rd)
  cor.test(fmri_vals$vmpfc_rd,fmri_vals$vs_rd)
  
  # questionnaire correlations
  cor.test(Q_results$AMI,fmri_vals$vmpfc_cd[1:n-1])
  cor.test(Q_results$AMI,fmri_vals$vmpfc_rd[1:n-1])
  plot(Q_results$AMI,fmri_vals$vmpfc_rd[1:n-1])
  cor.test(Q_results$AMI,fmri_vals$vs_rd[1:n-1])
  cor.test(Q_results$AMI,fmri_vals$vmpfc_m[1:n-1])
  cor.test(Q_results$HADS,fmri_vals$vmpfc_cd[1:n-1])
  plot(Q_results$HADS,fmri_vals$vmpfc_cd[1:n-1])
  cor.test(Q_results$HADS,fmri_vals$vmpfc_rd[1:n-1])
  cor.test(Q_results$HADS,fmri_vals$vs_rd[1:n-1])
  cor.test(Q_results$HADS,fmri_vals$vmpfc_m[1:n-1])
  cor.test(Q_results$GRIT,fmri_vals$vmpfc_cd[1:n-1])
  cor.test(Q_results$GRIT,fmri_vals$vmpfc_rd[1:n-1])
  cor.test(Q_results$GRIT,fmri_vals$vs_rd[1:n-1])
  plot(Q_results$GRIT,fmri_vals$vs_rd[1:n-1])
  cor.test(Q_results$GRIT,fmri_vals$vmpfc_m[1:n-1])
  
  # patient vs control 
  CrawfordHowell(fmri_vals$left_vs_roi_rwd[22],fmri_vals$left_vs_roi_rwd[1:21])
  CrawfordHowell(fmri_vals$left_vs_roi_rwd_d[22],fmri_vals$left_vs_roi_rwd_d[1:21])
}
```
analysis of patient vs healthy
```{r}
### Histograms comparing mean activity in ROI for several key regressors, patient vs controls ###

# motivation fmri values 
patient_m <- as.data.frame(fmri_vals$vmpfc_m)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_m
patient_m <- cbind(patient_m,p)

m_hist <- ggplot(patient_m,aes(x=fmri_vals$vmpfc_m,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - Motivation',x='M - mean beta weight from vmpfc ROI',y='No. Participants') 

## now for cd in vmpfc
patient_vmpfc_cd <- as.data.frame(fmri_vals$vmpfc_cd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_vmpfc_cd
patient_vmpfc_cd <- cbind(patient_vmpfc_cd,p)

vmpfc_cd_hist <- ggplot(patient_vmpfc_cd,aes(x=fmri_vals$vmpfc_cd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.1,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - Chosen Difference',x='vmpfc_cd - mean beta weight from vmpfc ROI',y='No. Participants') 

## now for rd in vmpfc
patient_vmpfc_rd <- as.data.frame(fmri_vals$vmpfc_rd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_vmpfc_rd
patient_vmpfc_rd <- cbind(patient_vmpfc_rd,p)

vmpfc_rd_hist <- ggplot(patient_vmpfc_rd,aes(x=fmri_vals$vmpfc_rd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - reward difference',x='vmpfc_rd - mean beta weight from vmpfc ROI',y='No. Participants') 

## now for rd in vs
patient_vs_rd <- as.data.frame(fmri_vals$vs_rd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_vs_rd
patient_vs_rd <- cbind(patient_vs_rd,p)

vs_rd_hist <- ggplot(patient_vs_rd,aes(x=fmri_vals$vs_rd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - reward diff',x='vs_rd - mean beta weight from vs ROI',y='No. Participants') 

## now for cd in vs
patient_vs_cd <- as.data.frame(fmri_vals$vs_cd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_vs_cd
patient_vs_cd <- cbind(patient_vs_cd,p)

vs_cd_hist <- ggplot(patient_vs_cd,aes(x=fmri_vals$vs_cd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - chosen diff',x='vs_cd - mean beta weight from vs ROI',y='No. Participants')

## now for cd in dmpfc
patient_dmpfc_cd <- as.data.frame(fmri_vals$dmpfc_cd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

#bind to patient_dmpfc_cd
patient_dmpfc_cd <- cbind(patient_dmpfc_cd,p)

dmpfc_cd_hist <- ggplot(patient_dmpfc_cd,aes(x=fmri_vals$dmpfc_cd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - chosen diff',x='dmpfc_cd - mean beta weight from dmpfc ROI',y='No. Participants')

## now for cd in lfp
patient_lfp_cd <- as.data.frame(fmri_vals$lfp_cd)

# create vector of 0s but last value 1, indicates which ones is patient
p <- rep(0,22)
p[22]=1

# bind to patient_lfp_cd
patient_lfp_cd <- cbind(patient_lfp_cd,p)

lfp_hist <- ggplot(patient_lfp_cd,aes(x=fmri_vals$lfp_cd,group=p,fill=p,color=p)) +
  geom_histogram(binwidth = 0.05,colour='black') +
  theme_classic() +
  theme(legend.position = 'none',
        axis.text=element_text(size=14),
        axis.title=element_text(size=18),
        plot.title=element_text(size=18)) +
  labs(title='Patient vs Controls - chosen_diff',x='lfp_cd - mean beta weight from lfp ROI',y='No. Participants')


```
