---
title: "checks"
author: "RB"
date: "2023-09-25"
output: html_document
---

load packages
```{r}
library(tidyverse)
library(wesanderson)
library(gridExtra)
library(ggpubr)
library(rstan)
library(plotrix)
source('scripts/reverseFromZero.R')
```

load online or offline data (online = in scanner)
```{r}
online = 1
```

load in the data 
```{r}
# read in the cleaned data, made in script 1.preprocess_2c.rmd
if (online == 1)
{
  real_data = read.csv('raw_data/clean/real_data_online.csv')
  sim_data = read.csv('raw_data/clean/sim_data_online.csv')
  parameters <- read.csv('raw_data/clean/parameters_online.csv')
} else {
  real_data = read.csv('raw_data/clean/real_data_offline.csv')
  sim_data = read.csv('raw_data/clean/sim_data_offline.csv')
  parameters <- read.csv('raw_data/clean/parameters_offline.csv')
}

# read in cleaned questionnaire data
Q_results <- read.csv('questionnaire_results.csv')

# read parameters
fmri_vals <- read.csv('raw_data/clean/fmri_vals.csv')

# set variable to data for analysis
data <- real_data

```

### colineaity of chosen devaluation and chose devaluaed option during devaluation?
```{r}
colinear <- data %>%
  select(correct,chosen_dev,isPurseFull_1,isPurseFull_2) %>%
  mutate(devaluation=(isPurseFull_1+isPurseFull_2)) %>%
  mutate(correct=correct*devaluation) %>% # isolate correct answers only in dev
  mutate(incorrect=1-correct) %>% # now isolate incorrect choices during extinction
  mutate(incorrect=incorrect*devaluation) %>%
  filter(devaluation == 1) %>%
  rowid_to_column() 

colinear[is.na(colinear)] <- 0

for (i in colinear$rowid){
  if (colinear[i,]$chosen_dev < 0.001){
    colinear[i,]$chosen_dev = 0
  } else if (colinear[i,]$chosen_dev > 0.001){
    colinear[i,]$chosen_dev = 1
  }
}

comp <- colinear %>%
  select(incorrect,chosen_dev)

cor(colinear$correct,colinear$chosen_dev)
cor(colinear$incorrect,colinear$chosen_dev)
```

#### STAY SWITCH ANALYSIS ###
Did participants choose the same option or switch when faced with certain outcomes?

probability of switching after receiving a reward or receiving nothing
```{r}
## create the data frame 
switch_stay <- data %>%
  select(subject_nr,totalWin,chosenDice) %>%
  group_by(subject_nr) %>%
  rowid_to_column()

# create switch column
switch_stay["switch"] <- 1

# set switch = 1 for switch and switch = 0 for stay
for (i in switch_stay$rowid){
  if (switch_stay[i,]$rowid == 1) next
  if (switch_stay[i,]$chosenDice == switch_stay[i-1,]$chosenDice){
    switch_stay[i-1,]$switch = 0
  }
}

# did they switch after a win
win_switch <- switch_stay %>%
  filter(totalWin == 1) %>%
  group_by(subject_nr) %>%
  mutate(stay = 1-switch)

mean(win_switch$switch)

# did they switch after getting nothing
nil_switch <- switch_stay %>%
  filter(totalWin == 0) %>%
  group_by(subject_nr) %>%
  mutate(stay = 1-switch)

mean(nil_switch$switch)

switch_stay <- switch_stay %>%
  group_by(subject_nr) %>%
  mutate(stay = 1-switch)
```


plot the probability of switching as a box plot
```{r}
plotting_2 <- switch_stay %>%
  group_by(totalWin, subject_nr) %>%
  mutate(p = mean(stay)) 
plotting_2["previous_outcome"] <- 'nothing'

for (i in plotting_2$rowid){
  if (plotting_2[i,]$totalWin == 1){
    plotting_2[i,]$previous_outcome = 'win'
  }
}

box_2 <- ggplot(plotting_2, aes(x=previous_outcome, y=p, group=previous_outcome, fill=previous_outcome)) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 2)) +
  geom_boxplot() +
  theme(axis.line = element_line(color='black'),
        panel.background = element_blank()) +
  labs(title = "Probabilty of choosing the same dice after wins or receiving nothing", x="previous outcome", y="p(stay)") +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  theme(text = element_text(size=12)) 
```


switch stay on trial 2 of reinstatement after loss, win devalued, win valued
```{r}
# create the data frame for reinstatement states
reinstatement <- data %>%
  select(state,stateTime,stateName,totalWin,isWinPurseFull,chosenDice,subject_nr) %>%
  filter(stateName == 'reinstatement') %>%
  rowid_to_column()

# create the prev outcome column
reinstatement['prev_outcome'] <- 'win'

# set previous outcome to win loss or devalued depending on outcome
for (i in reinstatement$rowid){
  if (reinstatement[i,]$rowid == 1) next
  if (reinstatement[i-1,]$totalWin == 0)
    reinstatement[i,]$prev_outcome = 'loss'
  else if (reinstatement[i-1,]$totalWin == 1 && reinstatement[i-1,]$isWinPurseFull == 1){
    reinstatement[i,]$prev_outcome = 'devalued'
  }
}

#create the switch column
reinstatement['switch'] <- 0

# set switch to 1 when they switched
for (i in reinstatement$rowid){
  if (reinstatement[i,]$rowid == 1) next
  if (reinstatement[i,]$chosenDice != reinstatement[i-1,]$chosenDice){
    reinstatement[i,]$switch = 1
  }
}

# create another data frame which removes t=1 as we only care about t=2 now
reinstatement_1 <- reinstatement %>%
  group_by(subject_nr,prev_outcome) %>%
  mutate(p = mean(switch)) %>%
  distinct(subject_nr,p,prev_outcome) 

switch_averages <- reinstatement_1 %>%
  filter(prev_outcome=='win')

### now do the same again but for trial 3
reinstatement <- data %>%
  select(state,stateTime,stateName,totalWin,isWinPurseFull,chosenDice,subject_nr) %>%
  filter(stateName == 'reinstatement') %>%
  filter(stateTime == 2 | stateTime == 3) %>%
  rowid_to_column()

reinstatement['prev_outcome'] <- 'win'

for (i in reinstatement$rowid){
  if (reinstatement[i,]$rowid == 1) next
  if (reinstatement[i-1,]$totalWin == 0)
    reinstatement[i,]$prev_outcome = 'loss'
  else if (reinstatement[i-1,]$totalWin == 1 && reinstatement[i-1,]$isWinPurseFull == 1){
    reinstatement[i,]$prev_outcome = 'devalued'
  }
}

#switch?

reinstatement['switch'] <- 0

for (i in reinstatement$rowid){
  if (reinstatement[i,]$rowid == 1) next
  if (reinstatement[i,]$chosenDice != reinstatement[i-1,]$chosenDice){
    reinstatement[i,]$switch = 1
  }
}

reinstatement_2 <- reinstatement %>%
  group_by(subject_nr,stateTime,prev_outcome) %>%
  mutate(p = mean(switch)) %>%
  distinct(subject_nr,p,prev_outcome) %>%
  filter(!stateTime == 2)

## correlations 
  cor.test(switch_averages$p,vmpfc_cd)
  cor.test(switch_averages$p,vmpfc_rd)
  cor.test(switch_averages$p,vs_rd)
  cor.test(switch_averages$p,vmpfc_m)
  cor.test(switch_averages$p,dmpfc_cd)
```


create box plots with these switch probabilities for the 2nd and 3rd trial after reinstatment 
```{r}
box_4 <- ggplot(reinstatement_1, aes(x=prev_outcome, y=p, group=prev_outcome, fill=prev_outcome)) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 3)) +
  geom_boxplot() +
  geom_jitter(aes(y=p, x=prev_outcome, group=subject_nr), width=0.2, height = 0) +
  theme(axis.line = element_line(color='black'),
        legend.position = 'none', 
        panel.background = element_blank()) + 
  labs(title = "Probabilty of switching on trial 2 of reinstatement", x="previous outcome", y="p(switch)") +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  theme(text = element_text(size=12)) 

box_5 <- ggplot(reinstatement_2, aes(x=prev_outcome, y=p, group=prev_outcome, fill=prev_outcome)) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 3)) +
  geom_boxplot() +
  geom_jitter(aes(y=p, x=prev_outcome, group=subject_nr), width=0.2, height=0) +
  theme(axis.line = element_line(color='black'),
        panel.background = element_blank(),
        legend.position = 'none') +
  labs(title = "Probabilty of switching on trial 3 of reinstatement", x="previous outcome", y="p(switch)") +
  scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  theme(text = element_text(size=12)) 

grid.arrange(box_4,box_5,ncol=2)
```
